---
output: github_document
pagetitle: README
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# licorer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The **licorer** package is designed to aid in the reading and processing of LI-6800 data files. It uses the R package [**units**](https://CRAN.R-project.org/package=units) to ensure that values recorded with units have the correct scientific units.

## Get **licorer**

<!-- from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("licorer") -->
<!-- ``` -->

from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("muir-lab/licorer")
```

and load **licorer**

```r
library("licorer")
```

## Vignette

```{r child = 'vignettes/loading-data.Rmd'}
```

## Contributors

* [Daniel Trupp](https://github.com/DanielTrupp)
* [Chris Muir](https://github.com/cdmuir)

## Comments and contributions

Comments, criticisms, and contributions are welcomed!
GitHub issues are the preferred way to report bugs, ask questions, or request 
new features.
You can submit issues here:
https://github.com/muir-lab/licorer/issues

## Meta

* Please [report any issues or bugs](https://github.com/muir-lab/licorer/issues)
* License: MIT + file LICENSE
