---
title: "Working with Data in LiCor Objects"
author: "Daniel Trupp"
date: "June 23, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Data in LiCor Objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[UTF-8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will show you how to take subsets of LI-COR data using the **licorer** package.

## Load packages and example data

```{r setup}
library(licorer)
library(units)
library(stringr)
library(knitr)
library(kableExtra)

test_data <- licor(system.file("extdata", "2019-05-06-0740_trillium_ovatum", package = "licorer", mustWork = TRUE))

```

Being a child of the "tibble" class, the "licor" class has inherited methods
for performing operations on data and displaying the results, the "licorer" 
package provides a few extra useful methods. We'll start with loading the data.

## Creting a Subset of the Data

There are two options for subsetting the data, `licor_subset()` and `subset_logic()`. The
`licor_subset()` option takes the data and a vector of options to subset the data by (taken
from the "subset_options" function) and returns a tibble with only those variables.
We'll show this with the "GasEx" option.

```{r}
subset_options(test_data)

scroll_box(knitr::kable(licor_subset(test_data, "GasEx")[1:5, ]),
            height = "250px", width = "700px", 
            box_css = "border: 1px solid #ddd; padding: 5px; ", 
            extra_css = NULL, fixed_thead = TRUE)
```

Once again we limit the size of the data shown to conserve space. This method is 
data-destructive, as the returned tibble is no longer a "licor" object and
lacks the header data. It does, however, preserve units.

The `subset_logic()` function works similarly, except it returns a logical vector 
to apply to the columns. It is non-data-destructive, as it creates a boolean vector.
This vetor is input in the column space to show the desired columns. We'll show
this with the "Sys" and "FlrLS" options.

```{r}
choice_vec <- c("Sys", "FlrLS")

scroll_box(knitr::kable(test_data[1:5, subset_logic(test_data, choice_vec)]),
            height = "250px", width = "700px", 
            box_css = "border: 1px solid #ddd; padding: 5px; ", 
            extra_css = NULL, fixed_thead = TRUE)
```

## Get remarks from datasets that have them

Any remarks that are read in from the file are attached to the data as an attribute.
The `get_remarks()` method will return a tibble of the remarks, and the lines before and after the remark was made.

``` {r}
test_data <- licor(system.file("extdata", "2020-06-23_ECT_hib", package = "licorer", mustWork = TRUE))

get_remarks(test_data)
```
