---
title: "Reading raw LI-6800 files into R"
author: "Daniel Trupp"
date: "June 23, 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{loading-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[UTF-8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(licorer)
library(knitr)
library(kableExtra)

test_data <- read_li6800_raw(system.file("extdata", "2020-06-23_ECT_hib", 
                                         package = "licorer", mustWork = TRUE))

```

The **licorer** package streamlines reading and validating data from raw LI-6800 files in *R*. To read a raw LI-6800 data file, use the `licor()` method. This will create a `licor` object, which stores the data in a tibble data object with header data stored in attributes.  

```{r, eval = FALSE}
library(licorer)

test_data <- read_li6800_raw(system.file("extdata", "2020-06-23_ECT_hib", 
                                         package = "licorer", mustWork = TRUE))
```

```{r, eval = TRUE, echo = FALSE}
scroll_box(knitr::kable(test_data, caption = "Example"),
            height = "250px", width = "700px", 
            box_css = "border: 1px solid #ddd; padding: 5px; ", 
            extra_css = NULL, fixed_thead = TRUE)
```

### The Data

As you can see, there is a lot of data. There are `r nrow(test_data)` rows and `r ncol(test_data)` columns. 

All of the data is stored either as a unit (some have the "unitless" unit) or as an approprate class (character, hms, or date/time), which are automatically assigned and validated when the file is loaded, as they are read in from the file. We'll discuss user-defined variables below.

LI-COR groups variables into a few different categories which are stored as attributes of the data. This topic is covered in greater detail in the vignette on [data manipulation](data-manipulation.html) and by the `subset_options()` function described below.

```{r}
subset_options(test_data)
```

The data columns covered by each of these values are expressly stated in the LI-6800 manual "Summary of symbols" section (CDM: provide page # and edition, v1.4 in this case).  

### The Header

In addition to data, the `licor` class stores meta-data from the header of the LI-6800 data file. This information is accessible with the `header` attribute.

```{r}
attributes(test_data)$header[1:5]
```

Once again we are showing a subset of the list, as the actual list of meta-data
contains around 50 elements. 

The `licor_validator()` that ensures that the `licor` class was created correctly and that all variables have valid scientific units. It is automatically called during the `licor` method call, but not during the `read_li6800_raw()` method.

Currently, the data reader only supports reading from LI-6800 files, but we 
plan to add support for the LI-6400 as well.
